### 2019-11-20 Manuscript/Supplementary_Information.tex

@@

\setcounter{figure}{0}
\renewcommand{\thefigure}{SI\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{SI\arabic{table}}

@@

\begin{SCfigure}[0.6][h!]
\begin{center}
	\includegraphics[width=0.6\textwidth]{../Figures/regolith_schematic.pdf}
	\caption{A schematic representation of the silicate weathering component of GEOCLIM in a single profile at steady-state. A rock particle leaves the unweathered bedrock with production rate $P_{r}$, and transits vertically through a regolith of height $h$. Regolith production and physical erosion ($E_{p}$) are equal at steady-state. As the particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$) with the flux of Ca and Mg corresponding to that weathering rate multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).}
	\label{fig:regolith_schematic}
\end{center}
\end{SCfigure}

The silicate weathering component of the GEOCLIM model has been modified from the previously published version \citep{Godderis2017b}. The new component implements the model of \citet{Gabet2009a} for the development of a chemically weathered profile. We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock. In the model of \citet{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, or as a physically weathered particle once it reaches the top. We use the DynSoil implementation of the \citet{Gabet2009a} model, which integrates chemical weathering within the regolith \citep{West2012a}. The transient time-varying version of this regolith model is described by three equations:

@@

\noindent
$P_{0}$ is the `optimal' regolith production rate (m/yr), which is defined to be the regolith production rate when there is no overlying regolith. In Equation \ref{eq:6}, we parameterize this value using the model of \citet{Carretier2014a}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the ideal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K). $f(h)$ is the soil production function (unitless), which describes how regolith production decreases as the thickness of the regolith increases. It takes an exponential form as is commonly applied in the literature (e.g. \citealp{Gabet2009a}). In Equation \ref{eq:7}, we follow \citet{Heimsath1997a}, where $d_{0}$ is a reference regolith thickness (m).

@@

\noindent
$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 \citep{Maffre2018a}. This formulation is directly inspired by the stream power law \citep{Davy2000a}. This formulation and these exponent values are supported by compilations, but variability in the proportionality constant is difficult to capture at a global scale \citep{Lague2013a}.

@@

\noindent
Equation \ref{eq:9} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \citep{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff (yr/m).

@@

\begin{table}[h!]
\begin{center}
\resizebox{0.7\textwidth}{!}{
    \begin{tabular}{cclcl}
    \textbf{GLiM} & \textbf{GLiM} & \textbf{GLiM} & \textbf{GEOCLIM} & \textbf{GEOCLIM}\\
    \textbf{ID} & \textbf{code} & \textbf{classification} & \textbf{ID} & \textbf{classification}\\
    &&&& \\
    \hline
    &&&& \\
    1 & su & unconsolidated sediments & 6 & sediments \\
    2 & vb & basic volcanic rocks & 4 & mafics \\
    3 & ss & siliciclastic sedimentary rocks & 6 & sediments \\
    4 & pb & basic plutonic rocks & 4 & mafics \\
    5 & sm & mixed sedimentary rocks & 6 & sediments \\
    6 & sc & carbonate sedimentary rocks & 5 & carbonates \\
    7 & va & acid volcanic rocks & 2 & felsics \\
    8 & mt & metamorphics & 1 & metamorphics \\
    9 & pa & acid plutonic rocks & 2 & felsics \\
    10 & vi & intermediate volcanic rocks & 3 & intermediates \\
    11 & wb & water bodies & 0 & water/ice \\
    12 & py & pyroclastics & 2 & felsics \\
    13 & pi & intermediate plutonic rocks & 3 & intermediates \\
    14 & ev & evaporites & 5 & carbonates \\
    15 & nd & no data & 1 & metamorphics \\
    16 & ig & ice and glaciers & 0 & water/ice \\
    \end{tabular}
}
\vspace{10pt}
\caption{Grouping of 16 lithologic categories in GLiM \citep{Hartmann2012a} to 6 broader categories for GEOCLIM.}
\label{tab:GLiM_to_GEOCLIM}
\end{center}
\end{table}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/world_lithology.jpg}
    \caption{Distribution of lithologies at 0.1\degrees $\times$ 0.1\degrees resolution used in GEOCLIM.}
    \label{fig:world_lithology}
\end{figure}

@@

The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) of \citet{Hartmann2012a}. The raw data takes the form of polygon vectors, where each polygon is assigned one of 16 lithologic categories. We first group these 16 categories into 6 broader categories (metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment; Table \ref{tab:GLiM_to_GEOCLIM}). We then rasterize the polygon vectors to 0.1\degrees $\times$ 0.1\degrees resolution, where each pixel is assigned the lithologic category of the polygon that covers the greatest area in that pixel (i.e. the `mode lithology'; Fig. \ref{fig:world_lithology}). To improve the computing time of GEOCLIM, we decrease the resolution of the raster to 0.5\degrees $\times$ 0.5\degrees. To do so, a 3-dimensional 720 $\times$ 360 $\times$ 7 matrix is created, in which the fraction of each 0.5\degrees $\times$ 0.5\degrees pixel covered by each of the 6 lithologic categories (or water/ice) is captured by the extra dimension. In this way, we calculate an area-weighted mean Mg+Ca concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/W_vs_r2.jpg}
    \caption{\textbf{A)} Modeled global \COtwo consumption vs. the coefficient of determination ($r^{2}$) between modeled and data-constrained \COtwo consumption in each of the watersheds. Each point represents model output using one of the 93,600 parameter combinations (Table \ref{tab:parameter_combinations}). \textbf{B)} Same as A, but zoomed to the plotting space with positive coefficient of determination values. The black line represents the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a}. The black box represents the range of these estimates preferred by \citet{Gerlach2011a}.}
    \label{fig:W_vs_r2}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/r2_cross_plot.pdf}
    \caption{Modeled vs. data-constrained \COtwo consumption in watersheds around the world. \textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 5,381 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (i.e. the orange points in Figure \ref{fig:W_vs_r2}). Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 5,381 parameters combinations. The red dashed line is the 1:1 line, the blue dashed line is the 2:1 line, and the cyan dashed lines are the 5:1 and 10:1 lines. \textbf{B)} Same as A, but only showing the parameter combination that produced that highest $r^{2}$. The chemical weathering and regolith thickness maps of this parameter combination are shown in Figure \ref{fig:weathering_regolith_maps}.}
    \label{fig:r2_cross_plot}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/weathering_regolith_maps.jpg}
    \caption{Chemical weathering and regolith thickness maps associated with the GEOCLIM model run using the parameter combination that produced the highest $r^{2}$ (Fig. \ref{fig:r2_cross_plot}B). Parameter values used for this model run are: $k_{d}=5\times10^{-4}$, $k_{w}=1$, $\sigma=-0.4$, $k_{rp}=1\times10^{-2}$, metamorphic$_{Ca+Mg}=2500$~mol/m$^{3}$, sediment$_{Ca+Mg}=2000$~mol/m$^{3}$.}
    \label{fig:weathering_regolith_maps}
\end{figure}

Experimental determinations of the activation energy ($E_a$; Equations \ref{eq:6} and \ref{eq:9}) associated with the weathering of silicate minerals are variable \citep{Brantley2003a}. However, multiple efforts to invert for $E_a$ in basaltic watersheds with varying temperature have yielded values (41.6 $\pm$ 3.2~kJ/mol in \citealp{Li2016a}; 42.3~kJ/mol in \citealp{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5 $\pm$ 1.7~kJ/mol; \citealp{Knauss1993a}) and for labradorite (42.1~kJ/mol; \citealp{Carroll2005a}). We use the value of 42~kJ/mol in our model runs. While we implement lithology-dependent Ca+Mg concentration, our implementation does not include lithology-dependent kinetics of mineral dissolution. Relative to felsic lithologies, mafic lithologies contain a higher concentration of minerals with faster dissolution kinetics (e.g. plagioclase). However, Ca+Mg from both felsic and mafic lithologies is predominantly sourced from minerals with faster dissolution kinetics, and we therefore use the same chemical weathering formulation (including the same activation energy) across lithologies.

Within the equation that governs the physical erosion rate in GEOCLIM (Equation \ref{eq:8}), the values of the exponents ($m=0.5$ and $n=1$) are supported by compilations \citep{Lague2013a}. However, there is uncertainty in the value of the proportionality constant ($k_e$). We therefore tune this value such that the total erosion flux in the model under present-day slope and runoff conditions (see below) matches the total erosion flux estimated in \citet{Milliman2013a} ($20\times10^{12}$~kg/yr). Assuming that the density of eroded materials is 2500~kg/m$^{3}$, our tuned $k_e$ value is 0.0029110~m$^{1-m}$/yr$^{1-m}$.

@@

We compute spatially-resolved long-term \COtwo consumption (i.e. Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0; \citealp{Fekete1999a}), temperature (CRU TS v.4.03; \citealp{Harris2013a}), and slope (Shuttle Radar Topography Mission; \citealp{Farr2007a}) fields. As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of \citet{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of \citet{Moquet2011a, Moquet2016a, Moquet2018a}. The latter are nested watersheds, which requires upstream weathering fluxes to be subtracted from downstream fluxes. Watersheds for which this subtraction yields an aberrant value are not considered. 80 watersheds in total are used in this study. We calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins:

@@

Global CO$_{2}$ consumption calculated from these 5,381 parameter combinations ranges from 2.9$\times10^{12}$~mol/yr to 7.7$\times10^{12}$~mol/yr, with a mean of 5.7$\times10^{12}$~mol/yr. These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in \citet{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig. \ref{fig:W_vs_r2}B).

Each parameter combination predicts a different total \COtwo consumption that should match the \COtwo degassing at steady-state. Hence, in the non-calibration GEOCLIM experiments, each parameter combination is used with its corresponding steady-state \COtwo degassing. In these estimates we are not including the effects of reverse weathering \citep{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \citep{Isson2018a}. Formation of authigenic clays that scales with Ca and Mg concentration of riverine waters could decrease the \COtwo consumption associated with silicate weathering although the overall trend seen in this study would remain.

@@

For the climate model component of GEOCLIM, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments (\citealp{Delworth2006a, Delworth2006b}; available for download at \url{https://nomads.gfdl.noaa.gov/dods-data/gfdl_cm2_0/}) These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca. 1860 at 2.0\degrees $\times$ 2.5\degrees resolution. In the ``1860 control'' experiment, forcing agents representative of conditions ca. 1860 (including \COtwo, CH$_{4}$, N$_{2}$O, O$_{3}$, sulfates, carbon, dust, sea salt, solar irradiance, and the distribution of land cover types) are held constant for 500 years after reaching equilibrium. \pCOtwo in 1860 is assumed to be 286~ppm. In the ``+1\%/yr to 2$\times$'' experiment, initial conditions are taken from the ``1860 control'' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value. \pCOtwo is then held constant until the end of the 280 year experiment. All non-\COtwo forcing agents are held constant. The ``+1\%/yr to 4$\times$'' experiment is identical to the ``+1\%/yr to 2$\times$'' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value. \pCOtwo is then held constant for 160 years. We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively. Temperature and runoff fields associated with \pCOtwo values between the three modeled \pCOtwo levels (286, 572, and 1144~ppm) are obtained through linear interpolation between the model results.

Given that precipitation in SEAI is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water \citep{Donohoe2017a}, the current approach of using a climate model forced by modern paleogeography for all SEAI scenarios is a reasonable approximation for this region. However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable. The climate model also assumes the present-day extent of ice-sheets for all \pCOtwo levels. However, without ice-sheets, the global climate would be warmer for a given \pCOtwo (hysteresis loop; \citealp{Pollard2005b}), but the amount of emergent land would be lower. We expect this bias to be minimal, as all of our \pCOtwo estimates fall below the glaciation threshold for Antarctica.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/SEAI_fracs.jpg}
    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of the Southeast Asian islands (SEAI) used to force GEOCLIM. Only the 15~Ma scenarios are shown here, but the 10 and 5~Ma scenarios would be similar, using the 10 and 5~Ma shorelines shown in Figure 1 of the main text instead. Each column represents a tested scenario. Each row represents a lithologic category. Solid black lines show present-day shorelines.}
    \label{fig:SEAI_fracs}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Manuscript/Figures/scenario_pCO2_all.pdf}
    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text. For each of the 13 scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (see SI). The box encloses the middle 50\% of the \pCOtwo estimates (i.e. the interquartile range), and the notch represents the median with its 95\% confidence interval. The whiskers extend to the 2.5 and 97.5 percentile values. This figure is identical to Figure 2 in the main text, except for the addition of the scenarios used to test the sensitivity of the results to the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet.}
    \label{fig:scenario_pCO2_all}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\textwidth]{Manuscript/Figures/paleogeographic_reconstructions.pdf}
    \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleogeographic model of \citet{Matthews2016a}.}
    \label{fig:paleogeographic_reconstructions}
\end{figure}

>>>>

% this isn't working... come back to it later
%\usepackage{xcite}
%\externalcitedocument{Manuscript}

@@

\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}

@@

\begin{SCfigure}[0.6][h!]
\begin{center}
	\includegraphics[width=0.6\textwidth]{Figures/regolith_schematic.pdf}
	\caption{A schematic representation of the silicate weathering component of GEOCLIM in a single profile at steady-state. A rock particle leaves the unweathered bedrock with production rate $P_{r}$, and transits vertically through a regolith of height $h$. Regolith production and physical erosion ($E_{p}$) are equal at steady-state. As the particle transits upwards, some fraction of the primary phases ($x$) are chemically weathered ($W$) with the flux of Ca and Mg corresponding to that weathering rate multiplied by the concentration of Ca+Mg in unweathered bedrock ($\chi_{CaMg}$).}
	\label{fig:regolith_schematic}
\end{center}
\end{SCfigure}

The silicate weathering component of the GEOCLIM model has been modified from the previously published version \cite{Godderis2017b}. The new component implements the model of Gabet and Mudd (2009) \cite{Gabet2009a} for the development of a chemically weathered profile. We refer to this chemically weathered profile as regolith where the base of the regolith is unweathered bedrock. In the model of Gabet and Mudd (2009) \cite{Gabet2009a}, material enters the regolith and leaves either as a solute through chemical weathering of the material during its travel from the bedrock towards the surface, or as a physically weathered particle once it reaches the top. We use the DynSoil implementation of the Gabet and Mudd (2009) \cite{Gabet2009a} model, which integrates chemical weathering within the regolith \cite{West2012a}. The transient time-varying version of this regolith model is described by three equations:

@@

\noindent
$P_{0}$ is the `optimal' regolith production rate (m/yr), which is defined to be the regolith production rate when there is no overlying regolith. In Equation \ref{eq:6}, we parameterize this value using the model of Carretier et al. (2004) \cite{Carretier2014a}, where $k_{rp}$ is a proportionality constant (unitless), $q$ is the runoff (m/yr), $E_{a}$ is the activation energy (J/K/mol), $R$ is the ideal gas constant (J/mol), $T$ is the temperature (K), and $T_{0}$ is the reference temperature (K). $f(h)$ is the soil production function (unitless), which describes how regolith production decreases as the thickness of the regolith increases. It takes an exponential form as is commonly applied in the literature \cite{Gabet2009a}. In Equation \ref{eq:7}, we follow Heimsath et al. (1997) \cite{Heimsath1997a}, where $d_{0}$ is a reference regolith thickness (m).

@@

\noindent
$k_{e}$ is a proportionality constant ((m/yr)$^{1-m}$) and $m$ and $n$ are adjustable exponents that are kept as 0.5 and 1 \cite{Maffre2018a}. This formulation is directly inspired by the stream power law \cite{Davy2000a}. This formulation and these exponent values are supported by compilations, but variability in the proportionality constant is difficult to capture at a global scale \cite{Lague2013a}.

@@

\noindent
Equation \ref{eq:9} is an empirical simplification of mineral dissolution rates derived from kinetic theory and laboratory experiments \cite{West2012a}, where $k_{d}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff and temperature (yr$^{-1-\sigma}$), and $k_{w}$ is a proportionality constant that modifies the dependence of dissolution rate on runoff (yr/m).

@@

\begin{table}[h!]
\begin{center}
\resizebox{0.7\textwidth}{!}{
    \begin{tabular}{cclcl}
    \textbf{GLiM} & \textbf{GLiM} & \textbf{GLiM} & \textbf{GEOCLIM} & \textbf{GEOCLIM}\\
    \textbf{ID} & \textbf{code} & \textbf{classification} & \textbf{ID} & \textbf{classification}\\
    &&&& \\
    \hline
    &&&& \\
    1 & su & unconsolidated sediments & 6 & sediments \\
    2 & vb & basic volcanic rocks & 4 & mafics \\
    3 & ss & siliciclastic sedimentary rocks & 6 & sediments \\
    4 & pb & basic plutonic rocks & 4 & mafics \\
    5 & sm & mixed sedimentary rocks & 6 & sediments \\
    6 & sc & carbonate sedimentary rocks & 5 & carbonates \\
    7 & va & acid volcanic rocks & 2 & felsics \\
    8 & mt & metamorphics & 1 & metamorphics \\
    9 & pa & acid plutonic rocks & 2 & felsics \\
    10 & vi & intermediate volcanic rocks & 3 & intermediates \\
    11 & wb & water bodies & 0 & water/ice \\
    12 & py & pyroclastics & 2 & felsics \\
    13 & pi & intermediate plutonic rocks & 3 & intermediates \\
    14 & ev & evaporites & 5 & carbonates \\
    15 & nd & no data & 1 & metamorphics \\
    16 & ig & ice and glaciers & 0 & water/ice \\
    \end{tabular}
}
\vspace{10pt}
\caption{Grouping of 16 lithologic categories in GLiM \cite{Hartmann2012a} to 6 broader categories for GEOCLIM.}
\label{tab:GLiM_to_GEOCLIM}
\end{center}
\end{table}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Figures/world_lithology.jpg}
    \caption{Distribution of lithologies at 0.1\degrees $\times$ 0.1\degrees resolution used in GEOCLIM.}
    \label{fig:world_lithology}
\end{figure}

@@

The spatial distribution of lithologies is sourced from the Global Lithologic Map (GLiM) of Hartmann and Moosdorf (2012) \cite{Hartmann2012a}. The raw data takes the form of polygon vectors, where each polygon is assigned one of 16 lithologic categories. We first group these 16 categories into 6 broader categories (metamorphic, felsic, intermediate, mafic, carbonate, and siliciclastic sediment; Table \ref{tab:GLiM_to_GEOCLIM}). We then rasterize the polygon vectors to 0.1\degrees $\times$ 0.1\degrees resolution, where each pixel is assigned the lithologic category of the polygon that covers the greatest area in that pixel (i.e. the `mode lithology'; Fig. \ref{fig:world_lithology}). To improve the computing time of GEOCLIM, we decrease the resolution of the raster to 0.5\degrees $\times$ 0.5\degrees. To do so, a 3-dimensional 720 $\times$ 360 $\times$ 7 matrix is created, in which the fraction of each 0.5\degrees $\times$ 0.5\degrees pixel covered by each of the 6 lithologic categories (or water/ice) is captured by the extra dimension. In this way, we calculate an area-weighted mean Mg+Ca concentration of the surface in each 0.5\degrees $\times$ 0.5\degrees pixel.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Figures/W_vs_r2.jpg}
    \caption{\textbf{A)} Modeled global \COtwo consumption vs. the coefficient of determination ($r^{2}$) between modeled and data-constrained \COtwo consumption in each of the watersheds. Each point represents model output using one of the 93,600 parameter combinations (Table \ref{tab:parameter_combinations}). \textbf{B)} Same as A, but zoomed to the plotting space with positive coefficient of determination values. The black line represents the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in Gerlach (2011) \cite{Gerlach2011a}. The black box represents the range of these estimates preferred by Gerlach (2011) \cite{Gerlach2011a}.}
    \label{fig:W_vs_r2}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Figures/r2_cross_plot.pdf}
    \caption{Modeled vs. data-constrained \COtwo consumption in watersheds around the world. \textbf{A)} Each point represents a single watershed, and the y-value of the point shows the mean value of the 5,381 parameter combinations that produce individual watershed \COtwo consumption fluxes that approximate those estimated in the literature for the present-day (i.e. the orange points in Figure \ref{fig:W_vs_r2}). Whiskers extend to the minimum and maximum modeled watershed \COtwo consumption fluxes for the 5,381 parameters combinations. The red dashed line is the 1:1 line, the blue dashed line is the 2:1 line, and the cyan dashed lines are the 5:1 and 10:1 lines. \textbf{B)} Same as A, but only showing the parameter combination that produced that highest $r^{2}$. The chemical weathering and regolith thickness maps of this parameter combination are shown in Figure \ref{fig:weathering_regolith_maps}.}
    \label{fig:r2_cross_plot}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Figures/weathering_regolith_maps.jpg}
    \caption{Chemical weathering and regolith thickness maps associated with the GEOCLIM model run using the parameter combination that produced the highest $r^{2}$ (Fig. \ref{fig:r2_cross_plot}B). Parameter values used for this model run are: $k_{d}=5\times10^{-4}$, $k_{w}=1$, $\sigma=-0.4$, $k_{rp}=1\times10^{-2}$, metamorphic$_{Ca+Mg}=2500$~mol/m$^{3}$, sediment$_{Ca+Mg}=2000$~mol/m$^{3}$.}
    \label{fig:weathering_regolith_maps}
\end{figure}

Experimental determinations of the activation energy ($E_a$; Equations \ref{eq:6} and \ref{eq:9}) associated with the weathering of silicate minerals are variable \cite{Brantley2003a}. However, multiple efforts to invert for $E_a$ in basaltic watersheds with varying temperature have yielded values (41.6 $\pm$ 3.2~kJ/mol \cite{Li2016a}; 42.3~kJ/mol \cite{Dessert2001a}) that are consistent with the lower end of activation energies of Ca+Mg bearing minerals in laboratory experiments such as that for diopside (40.5 $\pm$ 1.7~kJ/mol \cite{Knauss1993a}) and for labradorite (42.1~kJ/mol \cite{Carroll2005a}). We use the value of 42~kJ/mol in our model runs. While we implement lithology-dependent Ca+Mg concentration, our implementation does not include lithology-dependent kinetics of mineral dissolution. Relative to felsic lithologies, mafic lithologies contain a higher concentration of minerals with faster dissolution kinetics (e.g. plagioclase). However, Ca+Mg from both felsic and mafic lithologies is predominantly sourced from minerals with faster dissolution kinetics, and we therefore use the same chemical weathering formulation (including the same activation energy) across lithologies.

Within the equation that governs the physical erosion rate in GEOCLIM (Equation \ref{eq:8}), the values of the exponents ($m=0.5$ and $n=1$) are supported by compilations \cite{Lague2013a}. However, there is uncertainty in the value of the proportionality constant ($k_e$). We therefore tune this value such that the total erosion flux in the model under present-day slope and runoff conditions (see below) matches the total erosion flux estimated in Milliman and Farnsworth (2013) \cite{Milliman2013a} ($20\times10^{12}$~kg/yr). Assuming that the density of eroded materials is 2500~kg/m$^{3}$, our tuned $k_e$ value is 0.0029110~m$^{1-m}$/yr$^{1-m}$.

@@

We compute spatially-resolved long-term \COtwo consumption (i.e. Ca+Mg fluxes) using present-day runoff (UNH/GRDC Composite Runoff Fields V1.0 \cite{Fekete1999a}), temperature (CRU TS v.4.03 \cite{Harris2013a}), and slope (Shuttle Radar Topography Mission \cite{Farr2007a}) fields. As described in the main text, we then sum the computed \COtwo consumption over large-scale watersheds that appear in the global compilation of Gaillardet et al. (1999) \cite{Gaillardet1999a}, as well as smaller-scale watersheds of the Amazon Basin (HYBAM network) in the compilation of Moquet et al. (2011) \cite{Moquet2011a}, Moquet et al. (2016) \cite{Moquet2016a}, and Moquet et al. (2018) \cite{Moquet2018a}. The latter are nested watersheds, which requires upstream weathering fluxes to be subtracted from downstream fluxes. Watersheds for which this subtraction yields an aberrant value are not considered. 80 watersheds in total are used in this study. We calculate the coefficient of determination ($r^{2}$) between computed and measured \COtwo consumption in each of these basins:

@@

Global CO$_{2}$ consumption calculated from these 5,381 parameter combinations ranges from 2.9$\times10^{12}$~mol/yr to 7.7$\times10^{12}$~mol/yr, with a mean of 5.7$\times10^{12}$~mol/yr. These estimates fall within the range of those independently estimated in the literature for the present-day: the full range of estimates of the present-day non-anthropogenic global CO$_{2}$ emission rate compiled in Gerlach (2011) \cite{Gerlach2011a} is 3.0--10.0$\times10^{12}$~mol/yr, with the preferred range of those estimates being 3.4--5.9$\times10^{12}$~mol/yr (Fig. \ref{fig:W_vs_r2}B).

Each parameter combination predicts a different total \COtwo consumption that should match the \COtwo degassing at steady-state. Hence, in the non-calibration GEOCLIM experiments, each parameter combination is used with its corresponding steady-state \COtwo degassing. In these estimates we are not including the effects of reverse weathering \cite{Michalopoulos1995a} as it is not clear how it should be parameterized and it is interpreted to be a relatively minor flux in the Cenozoic \cite{Isson2018a}. Formation of authigenic clays that scales with Ca and Mg concentration of riverine waters could decrease the \COtwo consumption associated with silicate weathering although the overall trend seen in this study would remain.

@@

For the climate model component of GEOCLIM, we use temperature and runoff fields from a subset of the GFDL CM2.0 experiments \cite{Delworth2006a, Delworth2006b} (available for download at \url{https://nomads.gfdl.noaa.gov/dods-data/gfdl_cm2_0/}) These experiments were performed in order to explore the effect of various changes in forcing agents on climate since ca. 1860 at 2.0\degrees $\times$ 2.5\degrees resolution. In the ``1860 control'' experiment, forcing agents representative of conditions ca. 1860 (including \COtwo, CH$_{4}$, N$_{2}$O, O$_{3}$, sulfates, carbon, dust, sea salt, solar irradiance, and the distribution of land cover types) are held constant for 500 years after reaching equilibrium. \pCOtwo in 1860 is assumed to be 286~ppm. In the ``+1\%/yr to 2$\times$'' experiment, initial conditions are taken from the ``1860 control'' experiment, then \pCOtwo is prescribed to increase from 286~ppm at a compounded rate of +1\% per year for 70 years, when \pCOtwo reaches double (572~ppm) of the initial value. \pCOtwo is then held constant until the end of the 280 year experiment. All non-\COtwo forcing agents are held constant. The ``+1\%/yr to 4$\times$'' experiment is identical to the ``+1\%/yr to 2$\times$'' experiment, except that \pCOtwo is prescribed to increase for 140 years, when \pCOtwo reaches quadruple (1144~ppm) of the initial value. \pCOtwo is then held constant for 160 years. We take the mean of the last 100 years of each of these three experiments (when \pCOtwo is being held constant at its final level) to obtain temperature and runoff fields for 286, 572, and 1144~ppm respectively. Temperature and runoff fields associated with \pCOtwo values between the three modeled \pCOtwo levels (286, 572, and 1144~ppm) are obtained through linear interpolation between the model results.

Given that precipitation in SEAI is primarily driven by large-scale convective upwelling leading to high precipitation over both land and water \cite{Donohoe2017a}, the current approach of using a climate model forced by modern paleogeography for all SEAI scenarios is a reasonable approximation for this region. However, for paleogeographic change that would have very significantly modified precipitation over broad swathes of land such as growth of the Himalaya, this approach of using modern climate model results is not viable. The climate model also assumes the present-day extent of ice-sheets for all \pCOtwo levels. However, without ice-sheets, the global climate would be warmer for a given \pCOtwo (hysteresis loop \cite{Pollard2005b}), but the amount of emergent land would be lower. We expect this bias to be minimal, as all of our \pCOtwo estimates fall below the glaciation threshold for Antarctica.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Figures/SEAI_fracs.jpg}
    \caption{0.5\degrees $\times$ 0.5\degrees lithologic maps of the Southeast Asian islands (SEAI) used to force GEOCLIM. Only the 15~Ma scenarios are shown here, but the 10 and 5~Ma scenarios would be similar, using the 10 and 5~Ma shorelines shown in Figure 1 of the main text instead. Each column represents a tested scenario. Each row represents a lithologic category. Solid black lines show present-day shorelines.}
    \label{fig:SEAI_fracs}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{Figures/scenario_pCO2_all.pdf}
    \caption{Steady-state \pCOtwo estimates from GEOCLIM for the various scenarios discussed in the text. For each of the 13 scenarios, each point represents an estimate from one of the 5,381 unique parameter combinations that resulted in reasonable total global \COtwo consumption and most closely matched estimates of present-day \COtwo consumption in 80 watersheds around the world (see SI). The box encloses the middle 50\% of the \pCOtwo estimates (i.e. the interquartile range), and the notch represents the median with its 95\% confidence interval. The whiskers extend to the 2.5 and 97.5 percentile values. This figure is identical to Figure 2 in the main text, except for the addition of the scenarios used to test the sensitivity of the results to the inclusion of the Sunda Shelf and the removal of the Greenland ice sheet.}
    \label{fig:scenario_pCO2_all}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.7\textwidth]{Figures/paleogeographic_reconstructions.pdf}
    \caption{Paleogeographically-reconstructed paleoshorelines for SEAI at 5, 10, and 15~Ma, using the paleogeographic model of Matthews et al. (2016) \cite{Matthews2016a}.}
    \label{fig:paleogeographic_reconstructions}
\end{figure}
